---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nicknameR

<!-- badges: start -->
<!-- badges: end -->

nicknameR facilitates matching common English first names to common nicknames, a frequent problem when dealing with any name merging. The primary function returns a long or wide dataframe of names and nicknames. The function also allows on the fly formatting of the names.

The names data come from this GitHub repository: https://github.com/onyxrev/common_nickname_csv

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jmccrain/nicknameR")
```

## Usage



```{r}
library(nicknameR)

# standard long dataframe
names <- nicknamer()

dplyr::glimpse(names)

length(unique(names$name))
length(unique(names$nickname))
```

The default output is a long dataframe, with one row per name-nickname combination. In other words, both names and nicknames will be duplicated in each column, such as Alex being a nickname for Alexander or Alexandra.

Another option is to convert the output a wide format, using the `format` parameter. You can also transform the strings to either full uppercase or full lower case with `transform`.

```{r}
# wide dataframe transformed to lower case
names.wide <- nicknamer(format="wide", transform = "to_lower")

head(names.wide)
```

If you instead want to use the nicknames as the specific rows, you can set `reverse` to TRUE. Now there will be one row per *nickname*, so the name Alex will now have one row with two full names associated with it (Alexander and Alexandra). 

```{r}
# wide dataframe with nickname as the row
names.wide <- nicknamer(format="wide", transform = "to_lower", reverse=T)

```

Finally, the `opt_function` parameter allows you to pass any custom function to resulting dataframe. This transforms each name and nickname column.:


```{r, eval=F}
# optional function applied to names:
short.names <- nicknamer(
  format="wide",
  transform = "to_lower",
  opt_function = function(x) { str_sub(x, 1, 3) }
)
```

